---
title: "Study of Blast output for domains A and B of TopoVI"
author: "Leslie REGAD"
output:
  pdf_document:
    toc: no
    toc_depth: '3'
    number_sections: true
  html_document:
    code_folding: show
    df_print: paged
    fig_caption: yes
    highlight: pygments
    keep_md: no
    self_contained: yes
    theme: spacelab
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---


```{r echo=FALSE, warning=FALSE, message=FALSE}
#visualisation des tables 
library(DT)

#graphics
library(pheatmap)
#library(plotly)
library(FactoMineR)
library(igraph)
library(randomForest)
library(rpart)
library(rpart.plot)
library(rpart.plot)

options(width = 300)
# options(encoding = 'UTF-8')
knitr::opts_chunk$set(
  fig.width = 5.5, fig.height = 4, 
  fig.path = 'figures/asym_',
  fig.align = "center", 
  size = "tiny", 
  echo = TRUE, eval = TRUE, 
  warning = FALSE, message = FALSE, 
  results = TRUE, comment = "")
options(scipen = 12) ## Max number of digits for non-scientific notation
# knitr::asis_output("\\footnotesize")
```


# Data preparation

* Open file

Data were stored in the file `data/merge_060519.csv`. The two third columns contain sequence size for three subunits B and the two last columns contain sequence size for two  subunits A.


```{r}
fileIn <- read.table("data/merge_060519.csv", sep=";", header=T)
head(fileIn)
```

The file contains `r nrow(fileIn)` rows and `r ncol(fileIn)` columns.  


* Concatenate size of sequence for the three subunits B and for the two subunits A
```{r}
size.sB <- c(fileIn[,"Top6B_SACSH"], fileIn[,"Top6B_METMA"], fileIn[,"Top6B_HALMA"])
length(size.sB)

size.sA <- c(fileIn[,"Top6A_SACSH"],fileIn[,"Top6A_METMA"])
length(size.sA)
```


* We noted that some columns contains NA values. We removed these NA values.

```{r}
ind.supp.sB <- which(is.na(size.sB)==TRUE)
size.sB <- size.sB[-ind.supp.sB]

ind.supp.sA <- which(is.na(size.sA)==TRUE)
size.sA <- size.sA[-ind.supp.sA]
```

After removing NA value, we have :   
    + `r length(size.sB)` data for the subUnit B  
    + `r length(size.sA)` data for the subUnit A  
  

* concatenate these values in a list
```{r}
list.size <- vector("list", length=2)
names(list.size) <- c("sB","sA")

list.size[[1]] <-size.sB
list.size[[2]] <- size.sA

```


# Distribution of sequence size for subUnits A and B

We plot the distribution of sequence size for the two subunits using boxplot representation.

```{r}
boxplot(list.size, xlab="subunits", ylab="sequence size")
```

We noted that the two subunits do not have the same distribution.

* subunit B have more outliers than subunit A
* the sequence size associated to subunit B is more variable than this of subunit A

# Determination of descriptive statistic

## For subunit B : 

```{r boxplotB}
boxplot.sB <- boxplot(size.sB)
```


* average size :  `r round(mean(list.size[["sB"]]),2)` amino acids
* standard deviation : `r round(sd(list.size[["sB"]]),2)` amino acids  
$\rightarrow$ on average a sequence has `r round(mean(list.size[["sB"]]),2)` $\pm$ `r round(sd(list.size[["sB"]]),2)` amino acids

* minimum size :  `r round(min(list.size[["sB"]]),2)`  amino acids
* maximum size :  `r round(max(list.size[["sB"]]),2)`  amino acids  

* median size :  `r round(median(list.size[["sB"]]),2)`  amino acids  
$\rightarrow$ 50% of sequences have less than `r round(median(list.size[["sB"]]),2)`  amino acids

* first quartile size :  `r  quantile(list.size[["sB"]])[2]`  amino acids  
$\rightarrow$ 25% of sequences have less than `r quantile(list.size[["sB"]])[2]`  amino acids

* third quartile size :  `r  quantile(list.size[["sB"]])[4]`  amino acids  
$\rightarrow$ 75% of sequences have less than `r  quantile(list.size[["sB"]])[4]`  amino acids  



Recompute statistic parameters after removing outliers.  

* Outliers  = valeurs inférieures à la valeur pivot de gauche ou supérieures à la valeur pivot de droite  
* valeur pivot de gauche = Q1-1.5*(Q3-Q1) avec Q1 et Q3 = 1er et 3ème quartiles
* valeur pivot de droite = Q3 + 1.5*(Q3-Q1) avec Q1 et Q3 = 1er et 3ème quartiles  

```{r sB.ssOutliers}
size.sB.ssOutliers <- size.sB[which((size.sB> boxplot.sB$stats[1,1] ) &
                                      (size.sB< boxplot.sB$stats[5,1]))]
```
  
  
* average size after removing outliers :  `r round(mean( size.sB.ssOutliers))`  amino acids
* standard deviation after removing outliers :  `r round(sd( size.sB.ssOutliers))`  amino acids
* minimum size after removing outliers :  `r round(boxplot.sB$stats[1,1])`  amino acids
* maximun size after removing outliers :  `r round(boxplot.sB$stats[5,1])`  amino acids



## For subunit A : 

```{r boxplotA}
boxplot.sA <- boxplot(size.sA)
```


* average size :  `r round(mean(list.size[["sA"]]),2)` amino acids
* standard deviation : `r round(sd(list.size[["sA"]]),2)` amino acids  
$\rightarrow$ on average a sequence has `r round(mean(list.size[["sA"]]),2)` $\pm$ `r round(sd(list.size[["sA"]]),2)` amino acids

* minimum size :  `r round(min(list.size[["sA"]]),2)`  amino acids
* maximum size :  `r round(max(list.size[["sA"]]),2)`  amino acids  

* median size :  `r round(median(list.size[["sA"]]),2)`  amino acids  
$\rightarrow$ 50% of sequences have less than `r round(median(list.size[["sA"]]),2)`  amino acids

* first quartile size :  `r  quantile(list.size[["sA"]])[2]`  amino acids  
$\rightarrow$ 25% of sequences have less than `r quantile(list.size[["sA"]])[2]`  amino acids

* third quartile size :  `r  quantile(list.size[["sA"]])[4]`  amino acids  
$\rightarrow$ 75% of sequences have less than `r  quantile(list.size[["sA"]])[4]`  amino acids


Recompute statistic parameters after removing outliers.  

```{r sA.ssOutliers}
size.sA.ssOutliers <- size.sA[which((size.sA> boxplot.sA$stats[1,1] ) & 
                                      (size.sA< boxplot.sA$stats[5,1]))]
```

* average size after removing outliers :  `r round(mean( size.sA.ssOutliers))`  amino acids
* standard deviation after removing outliers :  `r round(sd( size.sA.ssOutliers))`  amino acids
* minimum size after removing outliers:  `r round(boxplot.sA$stats[1,1])`  amino acids
* maximun size after removing outliers:  `r round(boxplot.sA$stats[5,1])`  amino acids


# Comparison of the size variance for the two subunits

* Fisher test

We performed a Fisher test to compare the two variances

```{r var.test}
var.test(size.sB, size.sA)
```

Les conditions de validité du test (les échantillons doivent suivre une loi normale) n'est pas vérifiée.


* Wilcoxon test  
On va donc réaliser un test de wilcoxon sur les données suivantes : 

```{r}
data.sB <- abs(size.sB-mean(size.sB))
data.sA <- abs(size.sA-mean(size.sA))

wilcox.test(data.sB,data.sA)
```

The obtained p-value is smaller than 0.05.
the test is significant, thus we conclude that the variance of sequence size for subunit B is significant larger than sequence size for subunit A.



